\large{\bf Large effects and the infinitesimal model }

### 2023-07-07 writeup.tex
peter edits

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
\comment{TODO: Gregor suggests ``mendelian sampling term''}
What we will call the \emph{transmission noise}, $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}

@@

\comment{
    Proposal: describe model, pointing out it's the infinitesimal model with gaussian noise;
    say something about the Cauchy case (maintains diversity with blending inheritance?);
    and jump to equation (1);
    then present section 5.4 which we use to parameterize simulation.
}

As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.  The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution (\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$ is approximately $\frac{1}{2\pi\sigma^{2}} e^{-\frac{(z-m)^{2}}{2\sigma^{2}}}$ for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$.  Leaving aside -- as the Gaussian-population approximation frequently does -- the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if) deterministic mutation selection-balance is possible in an non-Gaussian model.

Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$ and is replaced by a new individual of phenotype
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution and $R$ is an independent draw from the transmission noise distribution (independent of $Z^{1}$ and $Z^{2}$).  Mutation selection balance demands that \comment{I'm struggling a bit to find a clear, short, and simple argument to jump directly to this point} that $\frac{Z^{1} +Z^{2}}{2} + R$ be equal in distribution to the selection-biased choice of $Z$, \ie for all bounded and continuous functions $f$,
\begin{equation} \label{eqn:MSB_condition}
    \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
\end{equation}
where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.

Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
\begin{align} \label{eqn:MSB_CF}
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
	\frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	&= \E\left[ e^{i\frac{theta}{2} Z^{1}}e^{i\frac{theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{\i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)	^{2} \Phi(\theta),
\end{align*}
where
\begin{equation}\label{eq:phi}
	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right]
\end{equation}
and
\begin{equation}\label{eq:Phi}
	\phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right]
\end{equation}
are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.

@@

First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta),
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$:
\[
	\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\varphi)(\theta).
\]
Iterating $F$ $n$ times, we get
\[
	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
\]
so that
\[
	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
\]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We will briefly explore what limiting values are possible -- and which limits are valid characteristic functions.  We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}. For the sake of simplicity, we will limit ourselves to the caae when both have well-posed limits independent of the other.

\subsubsection{Reproductive Terms}

First consider the limit
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
\]
Because $\psi(\theta)$ is the logarithm characteristic function, it must be continuous in some neighbourhood of, and vanish at 0, and thus $\alpha(\theta)$ as well.  Furthermore, for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.  We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.

Now, from the above, 
\[
	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
\]
as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
Now, if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
\[
	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1
\]
so $\gamma \geq 0$. If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and mean absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.

\paragraph{Lamarkian inheritance redux.} 
Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait \ie Lamarkian blending inheritance.  The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{Provine1971}.  The solution $\phi(\theta) = e^{im \theta -\gamma|\theta|}$ offers an amusing counter-factual: if a Lamarkian population could somehow achieved a Cauchy-distributed trait distribution, it would maintain that diversity indefinitely.

\paragraph{Other solutions?}
We would naturally like to know if these two exhaust the possible solutions.  Above, we exhausted all solutions for which $\alpha'(0+)$ exists.  However, if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist. This does not, however, in and of itself exclude $e^{\alpha(\theta)}$ from being a characteristic function: Bochner's theorem tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
\begin{enumerate}[(i)]
\item $\phi(\theta)$ is continuous,
\item $\phi(0) = 1$,
\item $\phi(-\theta) = \overline{\phi(\theta)}$, and
\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries $a_{ij} = \phi(\theta_{i}-\theta_{j})$
\end{enumerate}
One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.

\subsubsection{Noise Terms}

We now turn our attention to the noise terms.  A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.  Taylor's theorem then tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

>>>>

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
\comment{TODO: Gregor suggests ``Mendelian sampling term''}
What we will call the \emph{transmission noise}, $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}

@@

As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.  The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
Leaving aside
% -- as the Gaussian-population approximation frequently does --
the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if)
deterministic mutation-selection balance is possible in a non-Gaussian model
that retains independence of the Mendelian sampling terms.
(Indeed, above we have put substantial restrictions on what models of inheritance
might lead to such models.)
We might also ask whether such a model
might still reasonably approximate an additive model with non-Gaussian effects,
and for the purpose of investigating this with simulations,
some calculations are helpful.

Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
and is replaced by a the offspring of two randomly chosen individuals,
i.e., a new individual of phenotype
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
and $R$ is an independent draw from the Mendelian sampling distribution
(independent of $Z^{1}$ and $Z^{2}$). 
Mutation-selection balance demands that that
the distribution of individuals removed by death
and the distribution of individuals added by birth are equal,
\ie that $\frac{Z^{1} +Z^{2}}{2} + R$ is equal in distribution
to a $\mu$-weighted choice of $Z$.
% In other words, for all bounded and continuous functions $f$,
% \begin{equation} \label{eqn:MSB_condition}
%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
% \end{equation}
% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
% 
% Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
Taking the characteristic function of both these distributions,
deterministic mutation-selection balance is therefore equivalent to
\begin{align} \label{eqn:MSB_CF}
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
\end{align*}
where
$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
and
$ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
\ie
\begin{align*}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{align*}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions.  We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider to the case when both have well-posed limits independent of the other.

\paragraph{Reproductive Terms:}
First consider the limit
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
\]
Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanish at 0.
If $\phi$ is not differentiable at zero (and so the corresponding distribution
does not have a well-defined mean),
the limit does not exist.
On the other hand if $\psi$ is differentiable at zero,
then $\alpha$ is continuous there,
and for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.  We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.

Now, from the above, 
\[
	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
\]
as $k \to - \infty$ if the latter exists (\ie there is a unique right-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
Now, if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
\[
	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1
\]
so $\gamma \geq 0$. If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and mean absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.

\paragraph{Lamarkian inheritance redux.} 
Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait -- \ie Lamarkian blending inheritance.  The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{Provine1971}.  The solution $\phi(\theta) = e^{im \theta -\gamma|\theta|}$ offers an amusing counter-factual: if a Lamarkian population could somehow achieved a Cauchy-distributed trait distribution, it would maintain that diversity indefinitely.

% \paragraph{Other solutions?}
% We would naturally like to know if these two exhaust the possible solutions.  Above, we exhausted all solutions for which $\alpha'(0+)$ exists.  However, if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist. This does not, however, in and of itself exclude $e^{\alpha(\theta)}$ from being a characteristic function: Bochner's theorem tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
% \begin{enumerate}[(i)]
% \item $\phi(\theta)$ is continuous,
% \item $\phi(0) = 1$,
% \item $\phi(-\theta) = \overline{\phi(\theta)}$, and
% \item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries $a_{ij} = \phi(\theta_{i}-\theta_{j})$
% \end{enumerate}
% One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.

\paragraph{Noise Terms}
We now turn our attention to the noise terms.  A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.  Taylor's theorem then tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
